---
title: "turning point"
author: "liang kuang"
date: "April 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Turning points for oil price
In this R Markdown, it explores the Brend Spot Oil Price by trying to find the turning points from the time series. The first algorithm that experimented here is the most straigh-foward one, which defines a turning point as an event that has k consecutive change within the same direction and has momentum of at least above a threshold (In this case 5%).


```{r loadlibs}
#Load the libraries
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
library('ggplot2')
library('zoo')
library('dplyr')
library('readr')
library('dygraphs')
library('xts')
```


```{r loaddata}
# load the data
WTI_spot_price_IDDS <- read.csv("//was-2pbs01/Users/mfp53502/Projects/EEA/R/WTI_spot_price_IDDS.csv",header=TRUE,as.is=TRUE)
print(WTI_spot_price_IDDS$date[1])
WTI_spot_price_IDDS$date <- as.Date(WTI_spot_price_IDDS$date,"%m/%d/%Y")
str(WTI_spot_price_IDDS)
```

```{r find missing values}
#remove any NAs
WTI <- WTI_spot_price_IDDS[rowSums(is.na(WTI_spot_price_IDDS))==0,]
```

```{r plot}
# plot the WTI spot price
p1 <- ggplot() + geom_line(aes(y = datavalue, x = date,group = 1 ),size=0.5,
                               data = WTI)
p1 <- p1 + ggtitle('Brent Spot Oil Price')
p1
```
### Algorithm One
Three consecutive changes above 5%. (Monthly data from IDDS.)
```{r}
dIn <- WTI[,names(WTI)%in%c('date','datavalue')]
k = 3;
mag = 0.05;
numRec <- length(dIn[,2])
totalTP <- 0  # total number of turning points
TP_list <- list()
vout <- dIn[,2]
checkArray <- as.vector(matrix(0,nrow=k,ncol=1))
for(t in 1:(numRec-k)){
 #checkArray = as.vector(dIn[(t+1):(t+k),2] - dIn[t,2])
  checkArray = as.vector(diff(dIn[(t):(t+k),2])) # k consective same-direction change
  #print(checkArray)
 if(abs(sum(sign(checkArray)))==k & abs(checkArray[k]/dIn[t,2])>=mag) {
   # if duration >= k and absolute change >= mag
   TP_list <- append(TP_list,t)
 }else{
   vout <- NA
 }
}
TP_list = as.integer(as.vector(TP_list))
dOut <- data.frame(date=dIn[TP_list,1], value = dIn[TP_list,2])
dOutfull <- data.frame(data=dIn[,1], value = vout)
str(dOut)

```

### Experiment
Check the results by graphics and interactive plots using dygraphs. 
```{r plotresults}
dOut$date <- as.Date(dOut$date)
p1 <- p1 + geom_point(aes(x = date, y = value,group=1),size=1.5,color='red',
                           data = dOut)
p1
```
```{r datamanipulator}
#create an elemvated dummy time series for visualization only
dOut = mutate(dOut,value+5)
colnames(dOut) <- c('date','turning point','elevated') 

djoin <- merge(dIn,dOut, by=c('date'),all=TRUE)
str(djoin)
```

### Interactive exploration of the turning points
The javascript interactive graphic will enable us to zoom in and out the time series, also we can view the results by specifying the rolling window (rolling average smooth) in the lower-lest textbox. The bottom rangeselector also gives us a lot flexibility to view and validate the results. 

```{r interactivegraph, fig.width=9,fig.height=4}

djoin_xts <- xts(djoin[,2:3],order.by = djoin[,1])
dygraph(djoin_xts,main = 'Brent Oil Price Turning Points(experiment)') %>% dyRangeSelector() %>% dySeries('datavalue',label='Brent Oil Price',drawPoints = FALSE,color='black') %>% 
  dySeries('turning point', drawPoints = TRUE, color = 'red', pointSize = 2) %>%
  dyLegend(width=400,show='always', hideOnMouseOut = TRUE) %>% dyRoller(rollPeriod=1) %>% 
  dyHighlight(highlightCircleSize = 5,
              highlightSeriesBackgroundAlpha = 0.8,
              hideOnMouseOut = FALSE)%>%
  dyEvent("2008-07-01",'Financial Crisis',labelLoc = 'bottom') %>%
  dyEvent("2008-12-01",'Bottom', labelLoc = 'bottom')
#drawxts <- xts(WTI[,3],order.by = WTI[,2])
#dxts <- xts(dIn[,-1],order.by = dIn[,1]) 
#dygraph(drawxts) #%>% 
#dygraph(dxts,main='Brent Oil Price with Turning point (red)',xlab='Date') %>% dyRangeSelector()

```

### Conclusion
Based on the algorithm implemented here, it can capture the turning points pretty accurately. 

