---
title: "VAR Models - POC"
date: "10 April 2017"
output:
  html_document: default
  html_notebook: default
bibliography: library.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the packages
```{r}
suppressPackageStartupMessages({
library(vars)
library(urca)
library(knitr)
library(VARsignR)
library(pander)
library(MSBVAR)
})
```
###Data

I implemented a model for the global market of crude oil following the structure used by [@Killian2012]. The model includes 4 endogenous variables in logs: 

+ changes in global crude oil production: $\Delta oil.prod$, which captures aggregate supply shocks;

+ changes in an index of economic activity, measured by the global industrial production index: $\Delta global.ip$, assumed to capture aggregate demand shocks;

+ changes in real Brent oil prices, deflated using the US CPI: $\Delta roil.price$, designed to capture oil-market specific demand shocks; 

+ changes in global inventories of crude oil: $\Delta oil.inventory$, intended to capture speculative demand for oil. 

As exogenous variables I added seasonal dummies and an indicator for the recent declines in prices (dummy) that takes the value 1 during the 2008-09 recession and during the recent decline in oil prices between 2014 and 2016. 

```{r}
setwd("C:/Users/mfp43003/Desktop/EAA/data")

dt<-ts(xlsx::read.xlsx("EAA_Test_VAR.xlsx", sheetName = "VAR data", colIndex = 8:11,
                           startRow = 2,endRow = 242, header=T), start=c(1997,1),freq=12)
dummy<-ts(xlsx::read.xlsx("EAA_Test_VAR.xlsx", sheetName = "VAR data", colIndex = 12,
                              startRow = 2,endRow = 242, header=T), start=c(1997,1),freq=12)
```


####Plots of the data

```{r}
plot(dt, main="Endogenous variables included in the VAR")
```

Indicator of recent declines in oil prices has been included as an exogenous variable, together with seasonal dummies: 

```{r, echo=FALSE}
plot(dt[,"roil_price"],type="l",col=1, 
     ylab="Log (Brent)",xlab="Time",main="Recent sharp declines in Brent")
par(new=TRUE)
plot(dummy,type="l",col=3,xaxt="n",yaxt="n",xlab="",ylab="")
axis(4)
mtext("Declines in Brent",side=4,line=3, col=3)
legend("topleft",col=c(1,3),lty=1,legend=c("Log(Brent)","Declines indicator"))

```

####Stationarity 

The results of the unit root tests show that all the variables included in the analysis are integrated of order 1. In other words, these are first-difference stationary. Thus, in the VAR model I used the first difference of the data.

```{r}
uroots<-data.frame(dt[1:6,], row.names=c("Test value - levels","Critical Value at 5%","Test result - levels",
                                         "Test value - difference", "Critical value at 5%,", "Test results - difference"))

for(i in 1:ncol(dt)){
  uroots[1,i]<-round(ur.df(dt[,i], lags=12,type=c("drift"), selectlags = "BIC")@teststat[1,2],2)
  uroots[2,i]<-ur.df(dt[,i], lags=12,type=c("drift"), selectlags = "BIC")@cval[2,2]
  uroots[3,i]<-ifelse(uroots[1,i]>uroots[2,i],"Stationary","Unit Root")
  uroots[4,i]<-round(ur.df(diff(dt[,i]), lags=12, type="none",selectlags = "BIC")@teststat[1,1],2)
  uroots[5,i]<-ur.df(diff(dt[,i]), lags=12, type="none", selectlags = "BIC")@cval[1,2]
  uroots[6,i]<-ifelse(uroots[4,i]> uroots[5,i],"Stationary","Unit Root")
}
```


```{r echo = FALSE, results = 'asis'}
kable(t(uroots), caption = "Unit root test results")
```




###Simulations with impulse-response functions - unanticipated shocks to the economy 


#####Data for estimating the models - first difference
```{r}
d.dt<-diff(dt,lag=1)
d.dummy<-ts(cbind(d=dummy[2:nrow(dummy),]),start=c(1997,2),freq=12)
```


Recovering structural shocks from reduced-from VAR models is still an unresolved issue. In order to remain as agnostic as possible in terms of identifying restrictions, the method developed by [@Uhlig2005] of sign restrictions is tested here. Sign restrictions have become a popular alternative to the traditional approaches such as the Cholesky decomposition or zero restrictions, used to identify VAR systems [@Kilian2012; @Fry2011].

In this setting, prior beliefs about the sign of the impact of certain shocks can be used to identify structural shocks in VAR models [@Uhlig2005; @Canova2002]. The sign restrictions can be based on previous studies and formal models. 

Uhlig proposes the following procedure to partially identify a VAR model and and limit the number of a priori assumptions one needs to use in order to identify the structural shocks^[While this procedure aims to identify only one shock and not a complete identification of all the underlying structural shocks in the system, it can be extended to identify multiple simultaneous shocks.]: 

+ define sign restrictions for some of the variables in the system *on the impulse response functions* to a particular shock^[No zero restrictions are imposed and the sign restrictions are imposed on the *orthogonalized* impulse-response functions and not on the variable coefficients.];

+ pick a time period following the shock during which the sign restrictions are expected to hold - in this example the restrictions are assumed to hold for about 6 months; 

+ estimate a Bayesian VAR with a flat (i.e. uninformative) Normal inverted-Whishart prior - the model is estimated on log-differences of the data, except for interest rates; 

+ take draws from the posterior of the VAR parameters and compute impulse response functions; keep only the set of responses that satisfy the sign restrictions and repeat this step sufficiently often until we get a large sample of responses to perform inference analysis and construct error bands around the accepted draws. 


Under this identification procedure the models are set-identified as the sign restrictions do not generate a unique set of impulse responses [@Moon2011]. Thus, **while the structural shock of interest is identified, the model is unidentified under sign restrictions** [@Fry2011] as many different models can provide a similarly good fit to the data. 
Uhlig addresses this by proposing a penalty function that allows for a larger set of draws to be considered, even though they don't satisfy the sign restrictions. This approach imposes a penalty for violations of the sign restrictions but it doesn't reject the responses.

Robustness checks have been performed by identifying the economic shock using Uhlig's penalty function and [@Rubio-Ramirez2010] rejection method.^[These procedures are available in R in the package *VARsignR*.]


What are the effects of an improvement in the economic environment on the oil market? The only assumption made here is that a positive shock to the global industrial production index (the second variable in the data frame of endogenous variables) is believed to have a positive impact on oil production for a period of about six months. No constraints are imposed on responses of oil prices and oil inventories - these will be computed by the model. 

To identify the positive economic activity shock we only impose that a positive economic shock leads to an increase in global oil production. This assumption doesn't seem to be a very controversial consequence of an increase in global economic activity and it is supported by the data. 


```{r pander}
table <- rbind(Variables=colnames(dt), Shock=c("+", "+", "none","none"))
pandoc.table(table,style = "grid", caption = "Sign restrictons")
```


Impulse-response functions: the first element in the *constr* function below indicates the shock of interest in the model. Given the ordering of the variables in the data frame, the second variable represents the global industrial production index. 

The VAR model is estimated using the differenced data with two lags, no constant and assumes the sign restrictions last for 6 months following the shock. The impulse response functions are computed for a duration of 24 month following the positive shock to the economic activity. 

Mathematically, the second-order (2 lags) VAR model is given by: 

$y_{t} = c + A_{1}y_{t-1}+ A_{2}y_{t-2} + \epsilon_{t},     \epsilon_{t} \sim N(0, \Sigma)$

And the VAR in first differences is represented by: 

$\Delta y_{t} = B_{1}\Delta y_{t-1}+ B_{2}\Delta y_{t-2} + \epsilon_{t}$

Where $y_{t}$ represents the vector of endogenous variables: 

$$y_{t} = \begin{bmatrix}
\Delta oil.prod \\
\Delta global.ip \\
\Delta roil.price \\
\Delta oil.inventory
\end{bmatrix}
$$


The expanded equations are represented by: 

$\Delta oil.prod_{t} = a_{11}\Delta oil.prod_{t-1} + a_{12}\Delta oil.prod_{t-2} + a_{13}\Delta global.ip_{t-1} + a_{14}\Delta global.ip_{t-2} + a_{15}\Delta roil.price_{t-1} + a_{16}\Delta roil.price_{t-2} + a_{17}\Delta oil.inventory_{t-1} + a_{18}\Delta oil.inventory_{t-2}$
                       
$\Delta global.ip_{t} = a_{11}\Delta oil.prod_{t-1} + a_{12}\Delta oil.prod_{t-2} + a_{13}\Delta global.ip_{t-1} + a_{14}\Delta global.ip_{t-2} + a_{15}\Delta roil.price_{t-1} + a_{16}\Delta roil.price_{t-2} + a_{17}\Delta oil.inventory_{t-1} + a_{18}\Delta oil.inventory_{t-2}$
                       
$\Delta roil.price_{t} = a_{11}\Delta oil.prod_{t-1} + a_{12}\Delta oil.prod_{t-2} + a_{13}\Delta global.ip_{t-1} + a_{14}\Delta global.ip_{t-2} + a_{15}\Delta roil.price_{t-1} + a_{16}\Delta roil.price_{t-2} + a_{17}\Delta oil.inventory_{t-1} + a_{18}\Delta oil.inventory_{t-2}$
                       
                       
$\Delta oil.inventory_{t} = a_{11}\Delta oil.prod_{t-1} + a_{12}\Delta oil.prod_{t-2} + a_{13}\Delta global.ip_{t-1} + a_{14}\Delta global.ip_{t-2} + a_{15}\Delta roil.price_{t-1} + a_{16}\Delta roil.price_{t-2} + a_{17}\Delta oil.inventory_{t-1} + a_{18}\Delta oil.inventory_{t-2}$
                       


```{r, results="hide"}
constr <- c(+2,+1)

var2 <- uhlig.penalty(Y=d.dt, nlags=2, draws=2000, subdraws=1000, nkeep=1000, KMIN=1, KMAX=6, constrained=constr, constant=FALSE, crit=0.005, steps=24)

irfs1 <- var2$IRFS
vl <- c("Oil Production","Economic Activity","Brent (real price index)","Oil inventory")
```

The median of the impulse-response functions are plotted with 68% confidence bands: 

```{r}
irfplot(irfdraws=irfs1, type="median", labels=vl, save=FALSE, bands=c(0.16, 0.84), grid=TRUE, bw=FALSE)
```

An unanticipated positive economic shock (the magnitude equal one standard deviation of the industrial production index) leads to positive and statistically significant impact on oil prices and a negative effect on oil inventories. 


### Conditional forecasts with BVAR models 

Use the hard condition function *hc.forecast* from package **MSBVAR**. This function allows the forecasts of one or more of the endogenous variables in the VAR model to be constrained to special paths^[The unanticipated shocks simulated previously do not allow for conditions to be imposed on the endogenous variables.].



Fit a BVAR model with a flat prior

```{r}
bvar <- szbvar(dt, p=2, z=NULL, lambda0=1,
                         lambda1=0.1, lambda3=2, lambda4=0.5, lambda5=0,
                         mu5=3, mu6=3, nu=ncol(dt)+1, qm=12, prior=0,
                         posterior.fit=T)

```

Set up conditional forecast matrix conditions: forecasts should be computed for 24 months and the growth rate of the industrial production index is assumed to increase on average by one standard deviation of the historical data: 

```{r}
nsteps <- 24
ip.condition <- rep(6.8, nsteps)
yhat<-matrix(c(rep(0, nsteps),ip.condition, rep(0, nsteps*2)), ncol=4)
```

Set the random number seed so we can replicate the results and the conditional and the unconditional forecasts

```{r, results="hide"}
set.seed(11023)
conditional.forcs <- hc.forecast(bvar, yhat, nsteps, burnin=3000, gibbs=5000, exog=NULL)
unconditional.forcs <-uc.forecast(bvar, nsteps, burnin=3000, gibbs=5000, exog=NULL)
```

Set-up and plot the unconditional and conditional forecasts.  

```{r}
uc.flat <- NULL
hc.flat <- NULL
uc.flat$forecast <- unconditional.forcs$forecast[,,3:4]
hc.flat$forecast <- conditional.forcs$forecast[,,3:4]
```

Plot the conditional/unconditional forecasts and the actuals

```{r}
plot.hc <- function(fcasts1, varnames=NULL, start=c(0,1), 
                             freq=1, probs=c(0.05,0.95),
                             compare.level=NULL, ylab=NULL, ...)
  
{ 
  m <- dim(fcasts1$forecast)[3]
  h <- dim(fcasts1$forecast)[2]
  iters <- dim(fcasts1$forecast)[1]
  fcast1.summary <- array(apply(fcasts1$forecast, 3, forc.ecdf, probs=probs), c(h,3,m))

  par(las=1, mar=c(1,2,2.5,1))
  for(i in 1:m)
  { forc1.ci <- ts(fcast1.summary[,,i], start=start)
  
  ylim <- c(floor(min(c(forc1.ci))),
            ceiling(max(c(forc1.ci))))
  
  ts.plot(forc1.ci,  gpars=list(lty=c(1,1,1,2,2,2,4),
                     col=c("red","blue","blue"),
                     ylim=ylim, xlab="",axes=FALSE, ... ))
  axis(2,c(floor(min(c(forc1.ci))),
           ceiling(max(c(forc1.ci)))))
  mtext(varnames[i],side=3,line=1)
  
  box();
  if(i==1) { mtext(ylab, side=2, line=3, at=c(1.5*mean(ylim))) }
  abline(h=0)
  
  }
}

```

###References








